---
title: "Createing parameter table"
---

# Overview
## Objectives
+ To identify the which crops to be included in the analysis
+ Create a table that contains:
  + crop name of MIRCA-OS dataset
  + crop name of FAO report
  + parameter values (Ky, Kc) for each crop

## CROPWAT crop-parameter table

```{r}
# --- Load packages --- #
library(CropWat)

cp <- 
  load_crop_params() %>%
  data.table() # 55 crops

# --- Save --- #
fwrite(cp, here("Data/Raw/CropWat_crop_params_raw.csv"))
```

## Crops in CropGBWater dataset

```{r}
#/*--------------------------------*/
#' ## Explore which crops are covered in the dataset
#/*--------------------------------*/
ls_data_files <- 
  list.files(
    here("Data/Raw/CropGBWater/Data_S5_Y2020_WC_mm_gridded"),
    full.names = FALSE,
    pattern = "grid_avg.csv"
  )


ls_crops <-
  stringr::str_remove(ls_data_files, "Data_S5_Y2020_") %>%
  stringr::str_remove("_grid_avg.csv") %>%
  sort()


wheat_dt <- fread(here("Data/Raw/CropGBWater/Data_S5_Y2020_WC_mm_gridded/Data_S5_Y2020_Wheat_grid_avg.csv"))
```


# Constructing the crop-parameter table

We manually created a correspondence table between CropGBWater crop classes and FAO/CROPWAT crop classes. The file name is `CropGBWater_CROPWAT_correspondence.csv`. Then, we merge the crop correspondence table with the CROPWAT crop-parameter table to create a final crop-parameter table.




```{r}
crop_tbl <- fread("Data/Raw/CropGBWater_CROPWAT_correspondence.csv")

cropwat_tbl <- fread("Data/Raw/CropWat_crop_params_raw.csv")


# --- Merge --- #
Ky_Kc_tbl <- 
  merge(
    crop_tbl, cropwat_tbl, all.x = TRUE, by.x = "FAO_crop", by.y = "crop"
  ) %>%
  setnames(
    c("name", "Ytot"),
    c("FAO_crop_name,", "Ky")
  )

# --- Save --- #
fwrite(Ky_Kc_tbl, here("Data/Processed/CropGBWater_FAO_Ky_Kc.csv"))
```




