# Overview
## Objectives
+ For each crop $j$ and country $r$, create a table that contains:
  + $Q^{rain}_{jr}$: rainfed production quantity for each crop $j$ and country $r$.
  + $Q^{irr}_{jr}$: irrigated production quantity for each crop $j$ and country $r$.
  + $Q^{total}_{jr}$: total production quantity for each crop $j$ and country $r$.
+ Then, calculate the rainfed production share for each crop $j$ and country $r$ as follows:

$$
\text{Rainfed production share}_{jr} = \frac{Q^{rain}_{jr}}{Q^{total}_{jr}}
$$

## Data
+ SPAM 2020 v2.2 global production dataset for 2020. Downloaded from [link](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/SWPENT).
  + The unit is kg/ha


# Load data
```{r}
library(countrycode)
source(here("Scr/Agriculture/2-Irr_Code/DataPrep/0-functions-dataprep.R"))

# /*===== CropGBWater crop and SPAM crops correspondence =====*/
crop_gbwater_spam_tbl <- fread(here("Data/Raw/CropGBWater_SPAM_crop_correspondence.csv"))

#/*--------------------------------*/
#' ## SPAM 2020 data
#/*--------------------------------*/
spam_2020_total_raw <- 
  fread(here("Data/Raw/SPAM/2020/spam2020V2r2_global_P_TA.csv")) %>%
  .[, iso3 := countrycode(ADM0_NAME, origin = "country.name", destination = "iso3c")]

spam_2020_rain_raw <- 
  fread(here("Data/Raw/SPAM/2020//spam2020V2r2_global_P_TR.csv")) %>%
  .[, iso3 := countrycode(ADM0_NAME, origin = "country.name", destination = "iso3c")]

spam_2020_irr_raw  <- 
  fread(here("Data/Raw/SPAM/2020//spam2020V2r2_global_P_TI.csv")) %>%
  .[, iso3 := countrycode(ADM0_NAME, origin = "country.name", destination = "iso3c")]

#/*--------------------------------*/
#' ## CropGBWater Wheat variants
#/*--------------------------------*/
# ls_cols <- c("grid_code", "ADM0_NAME", "WHEA_A", "WHEA_I", "WHEA_R", "wheat_class", "WHEA_A")
ls_cols <- c("ADM0_NAME", "WHEA_A", "WHEA_I", "WHEA_R")

Wheat_spring_grid_avg <- 
  fread(here("Data/Processed/CropGBWater/wc/Wheat_spring_grid_avg.csv")) %>%
  .[, ..ls_cols] %>%
  .[, CropGBWater_crop := "Wheat_spring"]

Wheat_winter_grid_avg <-
  fread(here("Data/Processed/CropGBWater/wc/Wheat_winter_grid_avg.csv")) %>%
  .[, ..ls_cols] %>%
  .[, CropGBWater_crop := "Wheat_winter"]

Wheat_winter_fullcover_grid_avg <-
  fread(here("Data/Processed/CropGBWater/wc/Wheat_winter_fullcover_grid_avg.csv")) %>%
  .[, ..ls_cols] %>%
  .[, CropGBWater_crop := "Wheat_winter_fullcover"]


Wheat_grid_avg <- rbindlist(
  list(
    Wheat_spring_grid_avg,
    Wheat_winter_grid_avg,
    Wheat_winter_fullcover_grid_avg
  ),
  use.names = TRUE,
  fill = TRUE
)
```


## For wheat, calculate the share of each variant by technology type (irrigated, rainfed, total) for each country


```{r}
wheat_area_jr <- 
  Wheat_grid_avg[,
    .(
      wheat_area_total_ha = sum(WHEA_A, na.rm = TRUE),
      wheat_area_irr_ha   = sum(WHEA_I, na.rm = TRUE),
      wheat_area_rf_ha    = sum(WHEA_R, na.rm = TRUE)
    ),
    by = .(ADM0_NAME, CropGBWater_crop)
  ]

# Calculate the total wheat area for each country (denominator)
wheat_area_jr[
  ,
  `:=`(
    wheat_area_total_all_ha = sum(wheat_area_total_ha, na.rm = TRUE),
    wheat_area_irr_all_ha   = sum(wheat_area_irr_ha, na.rm = TRUE),
    wheat_area_rf_all_ha    = sum(wheat_area_rf_ha, na.rm = TRUE)
  ),
  by = ADM0_NAME
]


# Then, calculate the share of each wheat variant by technology type for each country
wheat_area_jr[
  ,
  `:=`(
    wheat_share_total_area = fifelse(
      wheat_area_total_all_ha > 0,
      wheat_area_total_ha / wheat_area_total_all_ha,
      NA_real_
    ),
    wheat_share_irr_area = fifelse(
      wheat_area_irr_all_ha > 0,
      wheat_area_irr_ha / wheat_area_irr_all_ha,
      NA_real_
    ),
    wheat_share_rf_area = fifelse(
      wheat_area_rf_all_ha > 0,
      wheat_area_rf_ha / wheat_area_rf_all_ha,
      NA_real_
    )
  )
]
```

## Derive country-level rainfed and irrigated production quantity

```{r}
ls_spam_crops <- unique(crop_gbwater_spam_tbl$SPAM_crop)
id_cols <- c("iso3", "ADM0_NAME")

spam_total_country <- 
  spam_2020_total_raw[,
    lapply(.SD, sum, na.rm = TRUE),
    by = .(iso3, ADM0_NAME),
    .SDcols = ls_spam_crops
  ] %>%
  melt(
    id.vars = id_cols,
    variable.name = "SPAM_crop",
    value.name = "Q_total"
  )

spam_rain_country <- 
  spam_2020_rain_raw[,
    lapply(.SD, sum, na.rm = TRUE),
    by = .(iso3, ADM0_NAME),
    .SDcols = ls_spam_crops
  ] %>%
  melt(
    id.vars = id_cols,
    variable.name = "SPAM_crop",
    value.name = "Q_rain"
  )

spam_irr_country <-
  spam_2020_irr_raw[,
    lapply(.SD, sum, na.rm = TRUE),
    by = .(iso3, ADM0_NAME),
    .SDcols = ls_spam_crops
  ] %>%
  melt(
    id.vars = id_cols,
    variable.name = "SPAM_crop",
    value.name = "Q_irr"
  )


#/*--------------------------------*/
#' ## Combine
#/*--------------------------------*/
spam_country_crop <- Reduce(
  function(x, y) {
    merge(
      x,
      y,
      by = c("iso3", "ADM0_NAME", "SPAM_crop"),
      all = TRUE
    )
  },
  list(
    spam_total_country,
    spam_rain_country,
    spam_irr_country
  )
)

spam_country_crop[
  ,
  `:=`(
    Q_total = fifelse(is.na(Q_total), 0, Q_total),
    Q_rain  = fifelse(is.na(Q_rain),  0, Q_rain),
    Q_irr   = fifelse(is.na(Q_irr),   0, Q_irr)
  )
]
```

## Direct mapping for non-wheat crops


```{r}
crop_gbwater_spam_nonwheat <- crop_gbwater_spam_tbl[SPAM_crop != "whea"]

spam_nonwheat_jr <- 
  merge(
    crop_gbwater_spam_nonwheat,
    spam_country_crop[SPAM_crop != "whea"],
    by = "SPAM_crop",
    all.x = TRUE,
    allow.cartesian = TRUE
  )

spam_nonwheat_jr <- 
  spam_nonwheat_jr[,
    .(
      iso3,
      ADM0_NAME,
      CropGBWater_crop,
      SPAM_crop,
      Q_rain,
      Q_irr,
      Q_total
    )
  ]
```


## Split aggregate SPAM wheat into your three wheat variants
```{r}
# This is country-level aggreated wheat production quantity
spam_wheat_jr <- 
  spam_country_crop[
    SPAM_crop == "whea",
    .(
      iso3,
      ADM0_NAME,
      SPAM_crop,
      Q_rain_whea  = Q_rain,
      Q_irr_whea   = Q_irr,
      Q_total_whea = Q_total
    )
  ]

wheat_split_jr <- 
  merge(
    wheat_area_jr,
    spam_wheat_jr,
    by = "ADM0_NAME",
    all.x = TRUE
  )

wheat_split_jr[,
  `:=`(
    Q_rain = fifelse(
      !is.na(wheat_share_rf_area),
      wheat_share_rf_area * Q_rain_whea,
      0
    ),
    Q_irr = fifelse(
      !is.na(wheat_share_irr_area),
      wheat_share_irr_area * Q_irr_whea,
      0
    )
  )
]

wheat_split_jr[, Q_total := Q_rain + Q_irr]

spam_wheat_variant_jr <- 
  wheat_split_jr[,
    .(
      iso3,
      ADM0_NAME,
      CropGBWater_crop,
      SPAM_crop = "whea",
      Q_rain,
      Q_irr,
      Q_total
    )
  ]
```

## Combine wheat and non-wheat data


```{r}
spam_cropgbwater_jr <- 
  rbindlist(
    list(
      spam_nonwheat_jr,
      spam_wheat_variant_jr
    ),
    use.names = TRUE,
    fill = TRUE
  )
```

## Calculate production share

```{r}
spam_cropgbwater_jr[,
  Q_total_for_share := Q_rain + Q_irr
]

spam_cropgbwater_jr[,
  `:=`(
    share_prod_rain = fifelse(
      Q_total_for_share > 0,
      Q_rain / Q_total_for_share,
      NA_real_
    ),
    share_prod_irr = fifelse(
      Q_total_for_share > 0,
      Q_irr / Q_total_for_share,
      NA_real_
    )
  )
]

spam_cropgbwater_jr[,
  share_prod_check := share_prod_rain + share_prod_irr
]


# === Save === #
fwrite(
  spam_cropgbwater_jr,
  here("Data/Processed/spam_cropgbwater_share.csv")
)
```

