rainfed_value_tbl
# Overview
## Objective
+ To create a table of $GPV_{jr}^{rain}$ and $GPV_{jr}^{irr}$ for each crop $j$ and country $r$.

## Data
+ FAOSTAT: Gross Production Value (GPV) by country and crop for the year 2019. Downloaded from [link](https://www.fao.org/faostat/en/#data/QV)
+ `spam_rainfed_share.csv`: A table of rainfed production share for each crop $j$ and country $r$ (created by `DataPrep/2-2-1-prep-spam.qmd`).
+ `s_jr_tbl.csv`: A table of $s_{rain, jr}$ for each crop $j$ and country $r$ (created by `DataPrep/2-1-lambda-calculation.qmd`).
  + This contains the information on the area of winter wheat and spring wheat separately, which is needed to allocate the total wheat GPV between winter wheat and spring wheat.


# NOTE:
+ For wheat, we need to split the total wheat GPV into winter wheat, spring wheat, and winter wheat full cover
+ For millet, we need to split the total millet GPV into pMillet and sMillet


# Data preparation


```{r}
library(countrycode)

#/*--------------------------------*/
#' ## Spam-CropGBWater share data
#/*--------------------------------*/
spam_cropgbwater_jr <- fread(here("Data/Processed/spam_cropgbwater_share.csv"))

#/*--------------------------------*/
#' ## FAOSTAT GPV data
#/*--------------------------------*/
gpv_raw <- 
  fread(here("Data/Raw/FAOSTAT/FAOSTAT_GPV_2020.csv")) %>%
  .[, iso3 := countrycode(`Area Code (M49)`, origin = "un", destination = "iso3c", custom_match = c("159" = "CHN"))] %>%
  # we keep only "China" and drop "China, mainland"
  .[!(iso3 == "CHN" & `Area` == "China, mainland")]
```


# FAOSTAT-to-CropGBWater mapping

```{r}
# unique(gpv_raw$Item)
# /*===== Correspondence table =====*/
crop_gbwater_faostat_gpv_tbl <- 
  fread(here("Data/Processed/CropGBWater_FAOSTAT_GPV_crop_correspondence.csv")) %>%
  .[, .(CropGBWater_crop, FAOSTAT_Item)] %>%
  unique()


#/*--------------------------------*/
#' ## Check correspondence table
#/*--------------------------------*/
missing_faostat_items <- 
  setdiff(
    unique(crop_gbwater_faostat_gpv_tbl$FAOSTAT_Item),
    unique(gpv_raw$Item)
  )

if (length(missing_faostat_items) > 0) {
  print(missing_faostat_items)
  stop("Some FAOSTAT items in the correspondence table are not found in gpv_raw.")
}

# Identify FAOSTAT items mapped to multiple CropGBWater crops
shared_faostat_items <- 
  crop_gbwater_faostat_gpv_tbl[
    ,
    .(n_crop = uniqueN(CropGBWater_crop)),
    by = FAOSTAT_Item
  ][
    n_crop > 1,
    FAOSTAT_Item
  ]

shared_faostat_items
# "Wheat" and "Millet"

#/*--------------------------------*/
#' ## Subset FAOSTAT GPV data to covered crops
#/*--------------------------------*/
gpv_item_tbl <- 
  merge(
    gpv_raw,
    crop_gbwater_faostat_gpv_tbl,
    by.x = "Item",
    by.y = "FAOSTAT_Item",
    all = FALSE
  )

#/*--------------------------------*/
#' ## Direct GPV table
#/*--------------------------------*/
# This is valid for one-to-one and many-to-one mappings.
# Example:
# Oranges + lemons + other citrus -> Citrus
gpv_direct_tbl <- 
  gpv_item_tbl[
    !Item %in% shared_faostat_items,
    .(
      GPV_total_jr = sum(as.numeric(Value), na.rm = TRUE),
      FAOSTAT_area = paste(sort(unique(Area)), collapse = "; "),
      FAOSTAT_items = paste(sort(unique(Item)), collapse = "; ")
    ),
    by = .(
      CropGBWater_crop,
      iso3
    )
  ][,GPV_unit := "1000 current US$"]

#/*--------------------------------*/
#' ## Shared FAOSTAT source GPV
#/*--------------------------------*/
# Important:
# Use gpv_raw here, not gpv_item_tbl.
# gpv_item_tbl already duplicates Wheat and Millet through the correspondence table.
gpv_shared_source_tbl <- 
  # gpv_raw[Item %in% shared_faostat_items,] %>%
  gpv_raw[
    Item %in% shared_faostat_items,
    .(
      GPV_source = sum(as.numeric(Value), na.rm = TRUE),
      FAOSTAT_area = paste(sort(unique(Area)), collapse = "; ")
    ),
    by = .(
      FAOSTAT_Item = Item,
      iso3
    )
  ][,GPV_unit := "1000 current US$"]
```

# Allocate shared FAOSTAT GPV to CropGBWater crops based on SPAM production shares

In this section, the goal is to allocate the shared FAOSTAT GPV for crops like "Wheat" and "Millet" to the corresponding CropGBWater crops (e.g., "Wheat_spring", "Wheat_winter", "pMillet", "sMillet") based on the production shares from SPAM.

+ 1. From `spam_cropgbwater_jr` and `shared_crop_map`, derive the total production quantity for each shared FAOSTAT item and country
  + this serves as the denominator for allocating GPV for each CropGBWater crop


```{r}
#/*--------------------------------*/
#' ## Allocate shared FAOSTAT GPV to CropGBWater crops
#/*--------------------------------*/
# Shared cases:
# Wheat  -> Wheat_spring, Wheat_winter, Wheat_winter_fullcover
# Millet -> pMillet, sMillet

shared_crop_map <- 
  crop_gbwater_faostat_gpv_tbl[FAOSTAT_Item %in% shared_faostat_items]

shared_spam_prod <- 
  merge(
    spam_cropgbwater_jr,
    shared_crop_map,
    by = "CropGBWater_crop",
    all = FALSE
  )

# Use SPAM production quantity to allocate FAOSTAT GPV across CropGBWater crops.
# First, calculate the total production quantity for each shared FAOSTAT item and country (denominator for allocation)
shared_spam_prod[,
  Q_total_shared := sum(Q_total_for_share, na.rm = TRUE),
  by = .(
    iso3,
    FAOSTAT_Item
  )
]

# Then, calculate the allocation share for each CropGBWater crop
shared_spam_prod[,
  GPV_alloc_share := fifelse(
    Q_total_shared > 0,
    Q_total_for_share / Q_total_shared,
    NA_real_
  )
]

# Merge the allocation share with the shared FAOSTAT GPV source table to calculate the allocated GPV for each CropGBWater crop
gpv_shared_alloc_tbl <- 
  merge(
    shared_spam_prod,
    gpv_shared_source_tbl,
    by = c("iso3", "FAOSTAT_Item"),
    all.x = TRUE
  ) %>%
  .[,GPV_total_jr := GPV_source * GPV_alloc_share]


gpv_shared_tbl <- 
  gpv_shared_alloc_tbl[
    ,
    .(
      CropGBWater_crop,
      iso3,
      GPV_total_jr,
      FAOSTAT_area,
      FAOSTAT_items = FAOSTAT_Item,
      GPV_unit
    )
  ]

# gpv_shared_source_tbl[iso3 == "AFG" & FAOSTAT_Item == "Millet"]
```

# Combine direct and shared GPV tables

```{r}
#/*--------------------------------*/
#' ## Combine direct and shared GPV tables
#/*--------------------------------*/
gpv_total_cropgbwater_jr <- 
  rbindlist(
    list(
      gpv_direct_tbl,
      gpv_shared_tbl
    ),
    use.names = TRUE,
    fill = TRUE
  )

#/*--------------------------------*/
#' ## Check duplicated crop-country rows
#/*--------------------------------*/
dup_gpv_check <- 
  gpv_total_cropgbwater_jr[, .N, by = .(CropGBWater_crop, iso3)] %>%
  .[N > 1]

if (nrow(dup_gpv_check) > 0) {
  print(dup_gpv_check)
  stop("Duplicated CropGBWater_crop-iso3 rows found in gpv_total_cropgbwater_jr.")
}

```

# Calculate rainfed and irrigated GPV using SPAM production shares
```{r}
#/*--------------------------------*/
#' ## Attach SPAM rainfed/irrigated production shares
#/*--------------------------------*/
spam_share_tbl <- copy(spam_cropgbwater_jr)

if (!"Q_total_for_share" %in% names(spam_share_tbl)) {
  spam_share_tbl[
    ,
    Q_total_for_share := Q_rain + Q_irr
  ]
}

if (!"share_prod_rain" %in% names(spam_share_tbl)) {
  spam_share_tbl[
    ,
    share_prod_rain := fifelse(
      Q_total_for_share > 0,
      Q_rain / Q_total_for_share,
      NA_real_
    )
  ]
}

if (!"share_prod_irr" %in% names(spam_share_tbl)) {
  spam_share_tbl[
    ,
    share_prod_irr := fifelse(
      Q_total_for_share > 0,
      Q_irr / Q_total_for_share,
      NA_real_
    )
  ]
}

spam_share_tbl <- spam_share_tbl[,
  .(
    CropGBWater_crop,
    iso3,
    Q_rain,
    Q_irr,
    Q_total_for_share,
    share_prod_rain,
    share_prod_irr
  )
]

#/*--------------------------------*/
#' ## Calculate rainfed and irrigated GPV
#/*--------------------------------*/
gpv_cropgbwater_jr <- merge(
  gpv_total_cropgbwater_jr,
  spam_share_tbl,
  by = c("CropGBWater_crop", "iso3"),
  all.x = TRUE
)

gpv_cropgbwater_jr[,
  `:=`(
    GPV_rain_jr = GPV_total_jr * share_prod_rain,
    GPV_irr_jr  = GPV_total_jr * share_prod_irr
  )
]

gpv_cropgbwater_jr[,
  GPV_check_diff := GPV_total_jr - (GPV_rain_jr + GPV_irr_jr)
]

#/*--------------------------------*/
#' ## Diagnostics
#/*--------------------------------*/
# Check rows with GPV but missing SPAM production shares
missing_share_check <- gpv_cropgbwater_jr[
  GPV_total_jr > 0 &
    (is.na(share_prod_rain) | is.na(share_prod_irr))
]

if (nrow(missing_share_check) > 0) {
  warning("Some crop-country GPV rows have missing SPAM rainfed/irrigated production shares.")
}

# Check GPV decomposition
gpv_decomp_check <- gpv_cropgbwater_jr[
  GPV_total_jr > 0 &
    !is.na(GPV_check_diff) &
    abs(GPV_check_diff) > 1e-6
]

if (nrow(gpv_decomp_check) > 0) {
  warning("Some crop-country rows do not satisfy GPV_total = GPV_rain + GPV_irr.")
}
```

# Add a status variable to describe why GPV_irr_jr and GPV_rain_jr are missing for some rows

```{r}
#/*--------------------------------*/
#' ## Record why rainfed/irrigated GPV is missing
#/*--------------------------------*/

gpv_cropgbwater_jr[
  ,
  GPV_split_status := fcase(
    GPV_total_jr > 0 &
      is.na(Q_total_for_share),
    "missing_SPAM_row",

    GPV_total_jr > 0 &
      !is.na(Q_total_for_share) &
      Q_total_for_share == 0,
    "zero_SPAM_production",

    GPV_total_jr > 0 &
      !is.na(share_prod_rain) &
      !is.na(share_prod_irr),
    "split_available",

    GPV_total_jr == 0,
    "zero_GPV",

    default = "other"
  )
]

gpv_cropgbwater_jr[
  ,
  GPV_split_note := fcase(
    GPV_split_status == "missing_SPAM_row",
    "FAOSTAT reports positive GPV, but no matching SPAM production-share row is available for this CropGBWater crop-country pair.",

    GPV_split_status == "zero_SPAM_production",
    "FAOSTAT reports positive GPV, but SPAM reports zero rainfed and irrigated production; rainfed/irrigated shares are undefined.",

    GPV_split_status == "split_available",
    "Rainfed and irrigated GPV are calculated using SPAM rainfed/irrigated production shares.",

    GPV_split_status == "zero_GPV",
    "FAOSTAT GPV is zero.",

    default = "Other unresolved case."
  )
]

gpv_cropgbwater_jr[
  GPV_split_status %in% c("missing_SPAM_row", "zero_SPAM_production"),
  `:=`(
    GPV_rain_jr = NA_real_,
    GPV_irr_jr  = NA_real_
  )
]
```

# Save the final table

```{r}
#/*--------------------------------*/
#' ## Save output
#/*--------------------------------*/
fwrite(
  gpv_cropgbwater_jr,
  here("Data/Processed/GPV_rainfed_irrigated_CropGBWater_2020.csv")
)
```