# Overview
## Objective
+ `Data_S5_Y2020_Wheat_grid_avg.csv` contains different types of wheat crops (i.e., winter wheat with full cover, winter wheat, and spring wheat). I want to create separate files for each wheat type, so that I can easily merge them with the corresponding crop calendar data in the future.


## Data
+ **CropGBWater (@Chukalla.etal2025)**, 2020 data, downloaded from https://zenodo.org/records/17059989

+ J_plant: Julian planting day, or day-of-year of planting date
+ Month: the calender month of planting
+ Day: the calendar month of planting


```{r}
# /*===========================================*/
#'=  Data Compilation (Don't run) =
# /*===========================================*/
wheat_calender_dt <-
  list.files(
    here("Data/Raw/CropGBWater/CropCalendarFolders/Wheat"),
    full.names = TRUE
  ) %>%
  lapply(., fread) %>%
  rbindlist()

# nrow(wheat_calender_dt) # 239138, each row is a unique grid information
# unique(wheat_calender_dt$Month)
# unique(wheat_calender_dt, by = c("Kc_ini", "Kc_mid", "Kc_end"))

# --- Save --- #
fwrite(wheat_calender_dt, here("Data/Processed/CropGBWater/wheat_cropcalender.csv"))
```

# Classification of crops

+ Kc_init values of 0.7, 0.4, and 0.3 correspond to FAO-W-WHEAT, FAO-W-WHEATF, and FAO-WHEAT, respectively. Use `Kc_ini` to classify wheat crops into three categories: FAO-W-WHEAT, FAO-W-WHEATF, and FAO-WHEAT.
  + FAO-W-WHEAT: Winter wheat, full-cover winter variant
  + FAO-W-WHEATF: Winter wheat
  + FAO-WHEAT: Spring wheat
+ To merge the wheat calendar data with the wheat water consumption data, we use `ID` in the calendar data and `ClimateID` in the water consumption data as the common key. 

```{r}
# --- Load wheat CropGBWater --- #
wheat_wc_dt <- fread(here("Data/Raw/CropGBWater/Data_S5_Y2020_WC_mm_gridded/Data_S5_Y2020_Wheat_grid_avg.csv"))
# nrow(wheat_dt) # 320978
# uniqueN(wheat_wc_dt$grid_code) # 320978

# --- Wheat calender crops --- #
wheat_calendar_dt <- fread(here("Data/Processed/CropGBWater/wheat_cropcalender.csv"))
# nrow(wheat_calendar_dt) # 239138
# uniqueN(wheat_calendar_dt$ID) # 239138

# setdiff(wheat_calendar_dt$ID, wheat_wc_dt$ClimateID) %>% length()
# intersect(wheat_calendar_dt$ID, wheat_wc_dt$ClimateID) %>% length() # all IDs in wheat_calendar_dt are in wheat_wc_dt

wheat_calendar_dt[
  , wheat_class := fcase(
    Kc_ini == 0.7, "winter_wheat_ff",
    Kc_ini == 0.4, "winter_wheat",
    Kc_ini == 0.3, "spring_wheat",
    default = NA_character_
  )
]

wheat_calendar_dt[
  , FAO_crop := fcase(
    Kc_ini == 0.7, "FAO-W-WHEAT",
    Kc_ini == 0.4, "FAO-W-WHEATF",
    Kc_ini == 0.3, "FAO-WHEAT",
    default = NA_character_
  )
]

#/*--------------------------------*/
#' ## Merge
#/*--------------------------------*/
wheat_calendar_key <- wheat_calendar_dt[
  , .(
    ClimateID = ID,
    wheat_class,
    FAO_crop,
    Kc_ini,
    Kc_mid,
    Kc_end,
    L_ini,
    L_dev,
    L_mid,
    L_late,
    J_plant,
    Month,
    Day,
    Totalperiod
  )
]

wheat_wc_split_dt <- merge(
  wheat_wc_dt,
  wheat_calendar_key,
  by = "ClimateID",
  all.x = TRUE
)

# nrow(wheat_wc_split_dt) == nrow(wheat_wc_dt) # TRUE

# --- Check --- #
wheat_wc_split_dt[
  , .N,
  by = .(
    valid_climate = ClimateID != -9999,
    matched_calendar = !is.na(wheat_class)
  )
]


# --- save --- #
fwrite(
  wheat_wc_split_dt[wheat_class == "winter_wheat_ff"],
  here("Data/Processed/CropGBWater/wc/Wheat_winter_fullcover_grid_avg.csv")
)

fwrite(
  wheat_wc_split_dt[wheat_class == "winter_wheat"],
  here("Data/Processed/CropGBWater/wc/Wheat_winter_grid_avg.csv")
)

fwrite(
  wheat_wc_split_dt[wheat_class == "spring_wheat"],
  here("Data/Processed/CropGBWater/wc/Wheat_spring_grid_avg.csv")
)
```