# Overview
## Objectives
For irrigated crops,:

+ construct daily crop water requirements (ETm_{jgtd}= Kc_{jh} * ET0_{gtd}), then aggregate ETm to crop-grid-month-stage
+ Allocate monthly CropGBWawter water consumption to stages
  + WCgc_{irr, jgm}: green water consumption on irrigated land
  + WCbl_{irr, jgm}: blue water consumption on irrigated land

## NOTE:
+ Crop-stage ETm is constructed using a representative 365-day daily ET0 profile averaged over 2018–2022. For growing seasons crossing December 31, days after year-end are mapped to the corresponding early-year day-of-year in the representative profile.

# Data loading

```{r}
# --- Load functions --- #
library(countrycode)
source(here("Scr/Agriculture/2-Irr_Code/DataPrep/0-functions-dataprep.R"))

# --- Suppress messages from data.table --- #
options(datatable.showProgress = FALSE)
options(datatable.verbose = FALSE)

# --- Crop Kc Ky Table --- #
crop_param_dt <- fread(here("Data/Processed/CropGBWater_FAO_Ky_Kc.csv"))

crop_param_long <- 
  crop_param_dt[,.(FAO_crop, CropGBWater_crop, crop_name, Yini, Ydev, Ymid, Yend)] %>%
  melt(
    id.vars = c("CropGBWater_crop", "FAO_crop", "crop_name"),
    measure.vars = patterns("^Y"),
    variable.name = "stage",
    value.name = "Ky"
  ) %>%
  .[, stage:= fcase(
    stage == "Yini", "ini",
    stage == "Ydev", "dev",
    stage == "Ymid", "mid",
    stage == "Yend", "late",
    default = NA_character_
  )]

ls_crops <- crop_param_dt$CropGBWater_crop

# --- Daily ET0 data --- #
et0_daily_dt <- fread(here("Data/Processed/CropGBWater/et0_daily_avg_2020.csv"))
```


# Function to calculate lambda for a given crop

```{r}
get_lambda <- function(crop_i){
  #/*--------------------------------*/
  #' ## Test Run
  # crop_i = ls_crops[1]
  # crop_i = "Banana" 
  # crop_i = "Cowpea"
  #/*--------------------------------*/

  message("Calculating lambda for crop: ", crop_i)

  # --- load wc data --- #
  wc_dt_crop_i <- fread(here("Data/Processed/CropGBWater/wc", paste0(crop_i, "_grid_avg.csv")))

  #/*--------------------------------*/
  #' ## Step 1: Calculate ETm_{jgtd}
  #/*--------------------------------*/
  # Crop water requirement accumulated over that month-stage portion
  etm_month_stage <- get_etm_month_stage(wc_dt_crop_i, et0_daily_dt)
   
  #/*--------------------------------*/
  #' ## Step 2: Allocate monthly CropGBWawter water consumption to stages
  #/*--------------------------------*/
  # (j, g, m, h) level
  wc_month_stage <- get_wc_month_stage(wc_dt_crop_i, etm_month_stage)

  # wc_month_stage[ClimateID == 212972]

  #/*--------------------------------*/
  #' ## Step 3: Convert mm to volume
  #/*--------------------------------*/
  # (j, g, m, h) level
  wc_month_stage_volume <- get_wc_stage_volume(wc_month_stage)

  #' + ETm_rf_stage_m3
  #' + ETm_ir_stage_m3
  #' + WC_bl_stage_m3
  #' + WC_gn_rf_stage_m3
  #' + WC_gn_ir_stage_m3 
   
  #/*--------------------------------*/
  #' ## Step 4: Aggregate grid-month-stage to grid-stage
  #/*--------------------------------*/
  # (j, r, h) level
  wc_stage_volume <-
    wc_month_stage_volume[,
    .(
      ETm_rf_jrgh_m3   = sum(ETm_rf_stage_m3, na.rm = TRUE),
      ETm_ir_jrgh_m3   = sum(ETm_ir_stage_m3, na.rm = TRUE),
      
      WC_gn_rf_jrgh_m3 = sum(WC_gn_rf_stage_m3, na.rm = TRUE),
      WC_gn_ir_jrgh_m3 = sum(WC_gn_ir_stage_m3, na.rm = TRUE),
      WC_bl_jrgh_m3    = sum(WC_bl_stage_m3, na.rm = TRUE),
      
      n_month_stage_records = .N,
      months_in_stage = paste(sort(unique(month_name)), collapse = ", ")
    ),
    by = c("ADM0_NAME", "grid_code", "ClimateID", "stage", "crosses_multiple_years")
  ]
  
  #/*--------------------------------*/
  #' ## Step 5: s_jrgmh calculation
  #/*--------------------------------*/
  # (j, g, h) level
  # For each grid-month-stage, calculate water satisfaction ratio for rainfed and irrigated land separately
  s_jgmh_tbl <- get_s_jrgh(wc_stage_volume)

  # --- check --- #
  # s_jgmh_tbl[!is.na(s_total_ir_check)]
 
  #/*--------------------------------*/
  #' ## Step 6: s_jrh calculation
  #/*--------------------------------*/
  # water satisfaction ratio for each crop-grid-month-stage
  s_jrh_tbl <- get_s_jrh(s_jgmh_tbl)

  # The kye outputs are:
  #' + s_gn_rf_jrh: green water satisfaction on rainfed land
  #' + s_gn_ir_jrh: green water satisfaction on irrigated land
  #' + s_bl_jrh: blue water satisfaction on irrigated land
  #' + s_total_ir_jrh: total water satisfaction on irrigated land (including both green and blue water)
 
  #/*--------------------------------*/
  #' ## Step 7: lambda calculation
  #/*--------------------------------*/
  # water satisfaction ratio for each crop-country

  # The key outputs are:
  #' + lambda_gn_rf: lambda for green water on rainfed land
  #' + lambda_gn_ir: lambda for green water on irrigated land
  #' + lambda_bl_ir: lambda for blue water on irrigated land
  #' + lambda_total_ir: lambda for total water on irrigated land (including both green
  crop_param_crop_i <- crop_param_long[CropGBWater_crop == crop_i]

  s_jrh_ky_tbl <- merge(s_jrh_tbl, crop_param_crop_i, by = "stage", all.x = TRUE)

  if (nrow(s_jrh_ky_tbl) != nrow(s_jrh_tbl)) {
    stop("Merge expanded or reduced rows for crop: ", crop_i)
  }

  if (s_jrh_ky_tbl[is.na(Ky), .N] > 0) {
    warning("Some stage rows do not have matched Ky values.")
  }

  lambda_jr_tbl <- 
    s_jrh_ky_tbl[,
      .(
        lambda_gn_rf = calc_lambda(s_gn_rf_jrh, Ky),
        lambda_gn_ir = calc_lambda(s_gn_ir_jrh, Ky),
        lambda_bl_ir = calc_lambda(s_bl_ir_jrh, Ky),
        lambda_total_ir = calc_lambda(s_total_ir_jrh, Ky),

        n_stage = uniqueN(stage),
        n_stage_with_Ky = sum(!is.na(Ky))
      ),
      by = .(
        ADM0_NAME,
        CropGBWater_crop,
        FAO_crop,
        crop_name
      )
    ]

  lambda_jr_tbl[,
    lambda_total_ir_check := lambda_gn_ir + lambda_bl_ir
  ][,
    lambda_total_ir_diff := lambda_total_ir - lambda_total_ir_check
  ]

  #/*--------------------------------*/
  #' ## Add some information
  #/*--------------------------------*/
  area_A_col <- grep("_A$", names(wc_month_stage), value = TRUE)
  area_I_col <- grep("_I$", names(wc_month_stage), value = TRUE)
  area_R_col <- grep("_R$", names(wc_month_stage), value = TRUE)


  keep_dt_grid <-
    wc_month_stage[,
      .(
        ADM0_NAME,
        grid_code,
        ClimateID,
        area_ha_A = get(area_A_col),
        area_ha_I = get(area_I_col),
        area_ha_R = get(area_R_col),    
        crosses_multiple_years
      )
    ] %>%
    unique(., by = c("ADM0_NAME", "grid_code"))

  keep_dt <- keep_dt_grid[,
    .(
      area_ha_A = sum(area_ha_A, na.rm = TRUE),
      area_ha_I = sum(area_ha_I, na.rm = TRUE),
      area_ha_R = sum(area_ha_R, na.rm = TRUE),
      
      n_grid = uniqueN(grid_code),
      n_ClimateID = uniqueN(ClimateID),
      
      any_crosses_multiple_years = any(crosses_multiple_years, na.rm = TRUE),
      
      area_ha_A_cross_year = sum(fifelse(crosses_multiple_years, area_ha_A, 0), na.rm = TRUE),
      area_ha_I_cross_year = sum(fifelse(crosses_multiple_years, area_ha_I, 0), na.rm = TRUE),
      area_ha_R_cross_year = sum(fifelse(crosses_multiple_years, area_ha_R, 0), na.rm = TRUE)
    ),
    by = .(ADM0_NAME)
  ]

  if (keep_dt[, anyDuplicated(ADM0_NAME)] > 0) {
    stop("Duplicate ADM0_NAME found in keep_dt for crop: ", crop_i)
  }

  # nrow(keep_dt) == nrow(lambda_jr_tbl) 

  lambda_jr_tbl_out <- merge(
    lambda_jr_tbl,
    keep_dt,
    by = "ADM0_NAME",
    all.x = TRUE
  )

  return(lambda_jr_tbl_out[])
}


# lambda_jr_tbl_maize <- get_lambda("Maize")
# lambda_jr_tbl_maize[ADM0_NAME %like% "United States", .(lambda_gn_rf, lambda_gn_ir, lambda_bl_ir, lambda_total_ir)]

# summary(lambda_jr_tbl_maize$lambda_gn_rf)
# summary(lambda_jr_tbl_maize$lambda_gn_ir)
# summary(lambda_jr_tbl_maize$lambda_bl_ir)
# summary(lambda_jr_tbl_maize$lambda_total_ir)


safe_get_lambda <- function(crop_i){
  warnings_i <- character()

  out <- tryCatch(
    withCallingHandlers(
      {
        lambda_i <- get_lambda(crop_i)

        lambda_i[
          ,
          requested_crop := crop_i
        ]

        list(
          success = TRUE,
          crop = crop_i,
          result = lambda_i,
          error = NA_character_,
          warnings = warnings_i
        )
      },
      warning = function(w){
        warnings_i <<- c(warnings_i, conditionMessage(w))
        invokeRestart("muffleWarning")
      }
    ),
    error = function(e){
      list(
        success = FALSE,
        crop = crop_i,
        result = NULL,
        error = conditionMessage(e),
        warnings = warnings_i
      )
    }
  )

  return(out)
}
```


# Run the function for all crops and combine results


```{r}
library(parallel)

lambda_results <- 
  mclapply(
    ls_crops,
    safe_get_lambda,
    mc.cores = 12
  )

# /*===== Error Table =====*/
lambda_error_tbl <- 
  rbindlist(
    lapply(lambda_results, function(x){
      if (!x$success) {
        data.table(
          crop = x$crop,
          error = x$error,
          warnings = paste(x$warnings, collapse = " | ")
        )
      } else {
        NULL
      }
    }),
    fill = TRUE
  )

lambda_warning_tbl <- 
  rbindlist(
    lapply(lambda_results, function(x){
      if (length(x$warnings) > 0) {
        data.table(
          crop = x$crop,
          warnings = paste(x$warnings, collapse = " | ")
        )
      } else {
        NULL
      }
    }),
    fill = TRUE
  )

# /*===== Results =====*/
lambda_jr_tbl <- 
  rbindlist(
    lapply(lambda_results, function(x){
      if (x$success) {
        x$result
      } else {
        NULL
      }
    }),
    fill = TRUE
  ) %>%
  .[, iso3 := countrycode(ADM0_NAME, origin = "country.name", destination = "iso3c")]

# setdiff(ls_crops, lambda_jr_tbl$CropGBWater_crop)

# lambda_jr_tbl[iso3 == "USA", .(CropGBWater_crop, lambda_gn_rf, lambda_gn_ir, lambda_bl_ir, lambda_total_ir)]


# /*===== Save =====*/
fwrite(
  lambda_jr_tbl,
  here("Data/Processed/lambda_jr_tbl_CropGBWater.csv")
)
```